---
title: |
  Vibe coding example\
  問いかけと出力記録
author: "Seiro Ito"
date: last-modified
date-format: "MMMM DD, YYYY, HH:MM ZZ"
execute:
  echo: false
  freeze: auto  # re-render only when source changes
engine: knitr
pdf-engine: lualatex
format:
  html:
    standalone: true
    embed-resources: true
    toc: true
    toc-location: left
    toc-depth: 1
    grid:
      margin-width: 350px
    linkcolor: "green"
    urlcolor: "#009900"
    template-partials: 
      - ../../assets/style/toc-slide.html
    include-after-body: ../../assets/style/toc-add.html
theme: [sky, ../../assets/style/sky-add.scss]
css:
  - ../../assets/style/my.css
  - https://cdn.jsdelivr.net/npm/tufte-css@1.8.0/tufte.min.css
filters:
  - ../../_extensions/tarleb/parse-latex/parse-latex.lua
  - ../../_extensions/seiro/tikzja/tikzja.lua
  - ../../assets/filters/ListArrow.lua
  - ../../_extensions/schochastics/nutshell/nutshell.lua
lightbox: auto
backgroundcolor: "#fffcf2"
title-block-banner: "#f0f3f5"
title-block-banner-color: "black"
doi-title: "doi.org/10.1086/734780"
callout-appearance: simple
callout-icon: false
bibliography: ../../seiro.bib
link-citations: true
reference-location: margin
citation-location: margin
---

# Meta prompt

Below is what I asked you in the terminal.

> I am a student who is analyzing data related to development and health. Find ViveCodingExample.qmd in this folder. Read the Procedure. Respond to my Instructions and render the file. I will update the file each time I receive your outputs written in the file. 

# Procedure

* I will add "Instructions" as you provide "Outputs." 
* You will add outputs right after each instruction. Also add "X+1th query" subsection, "Instructions", and "Outputs" when you respond to the X-th query.
* When you provide codes, they must be either in R or Python. R is the most preferred language. Use Python only if you cannot get the task done with R.
* Code blocks must not result in a horizontal scroll bar. Use linebreaks.
* Use a Tufte style css.
* Outputs must be succinct. Do not provide any analytical comments. I will do the interpretations. Provide only data related information that is useful in interpreting outputs, and the reason of your choices in creating outputs. 

# Instructions and outputs

## 1st query

### Instructions

* I am trying to find empirical relationship between child mortality and per capita health spending across countries. 
* Final output is a scatter plot with a line connecting each points, showing time series changes by country in a single figure.
* X axis is spendings, Y axis is mortality, and points are connected by a line for a same country. Use colour blind aware colour schemes, as in viridis.
* Find most suitable variables from WB data, WHO, DHS. 

### outputs

Variables (all from World Bank WDI, fetched with the R package `WDI`)

| Axis | Variable | WDI code | Unit | Original source |
|------|----------|----------|------|-----------------|
| Y | Under-5 mortality rate | `SH.DYN.MORT` | deaths per 1,000 live births | UN IGME (WHO/UNICEF/WB/UN) |
| X | Current health expenditure per capita | `SH.XPD.CHEX.PP.CD` | current int'l $, PPP | WHO Global Health Expenditure Database |

Reasons for choices

* Under-5 mortality: widest country × year coverage among child mortality series; WHO GHO publishes the same UN IGME estimates.
* Health spending in PPP dollars: removes cross-country price level differences, so X is comparable across countries.
* WDI rather than WHO GHO API: both variables in one call, same country codes and region labels.
* DHS not used: surveys are 5 or more years apart, country-specific years, and contain no health spending variable.
* Years 2000–2022: spending series starts in 2000.
* Both axes log10: spending ranges over about 100× across countries.
* Facets by WB region: 7 panels instead of one panel with over 200 colours.
* Legend off: colours only separate lines within a panel; viridis palette (colour-blind safe).
* Points connected in year order (`geom_path`), so each line traces a country over time.

```{r}
#| label: u5m-hexp
#| echo: true
#| cache: true
#| fig-width: 9
#| fig-height: 9
library(WDI)
library(dplyr)
library(ggplot2)
# WB API is slow: longer download timeout
old <- getOption("timeout")
options(timeout = 300)
d <- WDI(
  indicator = c(u5m  = "SH.DYN.MORT",
                hexp = "SH.XPD.CHEX.PP.CD"),
  start = 2000, end = 2022, extra = TRUE
) |>
  filter(region != "Aggregates",
         !is.na(region)) |>
  select(iso3c, country, region, year,
         u5m, hexp)
options(timeout = old)  # restore
dp <- d |>
  filter(!is.na(u5m), !is.na(hexp)) |>
  arrange(iso3c, year)
ggplot(dp, aes(hexp, u5m,
               colour = country,
               group = country)) +
  geom_path(alpha = .6) +
  geom_point(size = .6) +
  scale_x_log10() +
  scale_y_log10() +
  scale_colour_viridis_d() +
  facet_wrap(~ region, ncol = 2) +
  labs(x = paste("Health expenditure per",
                 "capita (PPP int'l $, log)"),
       y = paste("Under-5 mortality per",
                 "1,000 (log)")) +
  theme_minimal() +
  theme(legend.position = "none")
```

Data notes

```{r}
#| label: u5m-hexp-notes
#| echo: true
tibble(
  item = c("countries fetched",
           "countries plotted",
           "country-years plotted",
           "years",
           "missing u5m",
           "missing hexp"),
  value = c(n_distinct(d$iso3c),
            n_distinct(dp$iso3c),
            nrow(dp),
            paste(range(dp$year),
                  collapse = "-"),
            sum(is.na(d$u5m)),
            sum(is.na(d$hexp)))
) |> knitr::kable()
```

## 2nd query

### Instructions

### Outputs

